---
title: "PA1_template.Rmd"
author: "AbhishekShah"
date: "Sunday, December 20, 2015"
output: html_document
---

#Loading and preprocessing the data
##1. Load the data (i.e. read.csv())

```{r, echo=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

```{r}
if(!file.exists('activity.csv')){
    unzip('repdata-data-activity.zip')
}
activityData <- read.csv('activity.csv')
```

##2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r, echo=TRUE}
activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```

#What is mean total number of steps taken per day?
```{r, echo=TRUE}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```

##1. Make a histogram of the total number of steps taken each day
```{r}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

##2. Calculate and report the mean and median total number of steps taken per day

```{r, echo=TRUE}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```

- Mean:
```{r, echo=FALSE} 
stepsByDayMean
```
- Median:
```{r, echo=FALSE} 
stepsByDayMedian
```

#What is the average daily activity pattern?
```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=data.frame(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

##1. Make a time series plot
```{r}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```
##2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```
- Most Steps at:
```{r, echo=FALSE}
timeMostSteps
```

#Imputing missing values
##1. Calculate and report the total number of missing values in the dataset
```{r, echo=TRUE}
numMissingValues <- length(which(is.na(activityData$steps)))
```
- Number of missing values:
```{r, echo=FALSE}
numMissingValues
```

##2. Devise a strategy for filling in all of the missing values in the dataset.

##3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```

##4. Make a histogram of the total number of steps taken each day

```{r}
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```
##... and Calculate and report the mean and median total number of steps taken per day.

```{r}
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
- Mean (Imputed):
```{r}
stepsByDayMeanImputed
```
- Median (Imputed):
```{r}
stepsByDayMedianImputed
```

#Are there differences in activity patterns between weekdays and weekends?

##1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##2. Make a panel plot containing a time series plot
